Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire

### 2016-09-15 manuscript/supplementary.Rnw
More edits for GCB

\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{gcbresearcharticle} % Choose template

@@

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into naturally occuring inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. S\ref{figs8}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

>>>>

\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment. 
\pdfminorversion=4
\templatetype{gcbresearcharticle} % Choose template

@@

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into naturally occuring inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. \ref{figs8}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. \ref{figs5}-\ref{figs7}). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation \citep[equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C;][]{Brown:1989vx} influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:
